---
title: "Census API Data"
author: "Diversiteam (VAI)"
date: "11/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Choosing the data  

##### First, take a look at all of the available US Census APIs
``` {r}
# Load libraries
library(censusapi)
library(tidyr)
library(stringr)
library(dplyr)
library(tidyverse)

# Force read the R environment file
readRenviron("~/.Renviron")

# Create a new dataframe with the full list of available census APIs
apis <- listCensusApis()

# Take a look at the available APIs
# View(apis)
```

##### Balancing breadth and depth, we decided to use the 2018 ACS 5-year data  
Look at the metadata and make some decisions about what to keep (to-do offline).
``` {r}
# Create a new dataframe to store the metadata for the 2018 ACS 5-year dataset
acs_vars <- listCensusMetadata(
 name = "acs/acs5/profile",
 vintage = "2018",
 type = "variables"
)

# Take a look at the metadata dataframe
# View(acs_vars)

# Count the number of times '!!' shows up as a separator in the `label` field
max(str_count(acs_vars$label, "!!"))

# Split the `label` field into their own fields and keep the original `label` field
acs_vars <- acs_vars %>%
     separate(
     label,
     c('label_01', 'label_02', 'label_03', 'label_04', 'label_05', 'label_06', 'label_07', 'label_08'),
     sep = "!!",
     remove = FALSE,
     convert = FALSE
     )

# Order by the `label_02` column
acs_vars <- arrange(acs_vars, concept, label_01, label_02, label_03, label_04, label_05, label_06, label_07, label_08)
```

##### Write to a CSV for group perusal

``` {r}
write.csv(acs_vars, "data/acs_5y_2018_metadata.csv", row.names = FALSE)
```

##### Start looking at the fields in the ACS 2018 5-year data

Just a preliminary exploration.
``` {r}
# Take a look at the unique values in the concept field
acs_concepts <- unique(acs_vars$concept)
View(acs_concepts)

# Take a look at the unique values in the label_01 field
acs_label_01 <- unique(acs_vars$label_01)
View(acs_label_01)

# Take a look at the unique values in the label_02 field
acs_label_02 <- unique(acs_vars$label_02)
View(acs_label_02)
```

##### Select only the labels we need for the project

``` {r}
# Create a list of the label_02 values we want to filter on
label_02s <- c('DISABILITY STATUS OF THE CIVILIAN NONINSTITUTIONALIZED POPULATION',
'EDUCATIONAL ATTAINMENT',
'HISPANIC OR LATINO AND RACE',
'HOUSEHOLDS BY TYPE',
'LANGUAGE SPOKEN AT HOME',
'MARITAL STATUS',
'PERCENTAGE OF FAMILIES AND PEOPLE WHOSE INCOME IN THE PAST 12 MONTHS IS BELOW THE POVERTY LEVEL',
'SEX AND AGE',
'U.S. CITIZENSHIP STATUS',
'VETERAN STATUS')

# Create a new dataframe for only the variables we want and drop the columns we won't need
target_acs_vars <- acs_vars %>% 
        filter(label_02 %in% label_02s) %>% 
        filter(label_01 == "Estimate") %>% 
        filter(concept != "SELECTED SOCIAL CHARACTERISTICS IN PUERTO RICO") %>% 
        select(-one_of(c("label", "label_07", "label_08", "predicateType", "group", "limit", "attributes", "required")))

View(target_acs_vars)
# Did it work?
# View(target_acs_vars)
```

##### Get the data for Davidson County, TN

Link to the state / county codes here:
``` {r}

# Check to see what geo fields there are in the 2018 ACS 5-year dataset
listCensusMetadata(
        name = "acs/acs5/profile",
        vintage = "2018",
        type = "geography"
) # %>% # add me if you want to view it immediately
  #      View()

# Turn the name column from our target_acs_vars table into a list
target_names <- target_acs_vars$name

# Check to see if it worked
# View(target_names)

# Get the data we want from the census API
acs_data <- getCensus(
        name = "acs/acs5/profile",
        vintage = "2018",
        region = "county:037",
        regionin = "state:47",
        vars = target_names
)

# Pivot the dataframe so there aren't 195 columns
acs_data <- pivot_longer(
        acs_data,
        cols = starts_with("DP"),
        names_to = "name",
        values_to = "value"
)

# Join the dataframe to the metadata dataframe to get all of the label information
acs_data <-
        inner_join(acs_data, target_acs_vars, by = "name", copy = FALSE)

# Write it out to a csv
write.csv(acs_data, "data/acs_5y_2018_data.csv", row.names = FALSE)
```

